Neural models for computer-assisted musical orchestration: a preliminary study

### 2020-05-12 paper/latex/paper.tex
Revise data augmentation

In order to increase the variability of the samples, we used two methods of data augmentation on the samples.\\

The first one was done on the mixture of the samples, by slightly shifting the frequency of the signal. In practice, the sample rate used to compute the Mel spectrogram was modified to be within $5\%$ of the acual $44100Hz$.\\

The second one was done on the Mel spectrogram, by setting to 0 random columns and rows of the feature matrices. For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$ reset columns and $3.45$ reset rows. This method of data augmentation aimed to be more resilient to the possible errors of recording, due either to a setup that induces interferences and cancels specific frequences, or to a default in the microphone that fails to record a small temporal portion.

\section{Our model}
- deep learning trained for classification and then used for orchestraion
- changing number of instruments

- explain why we generated our data in the way we did
Since our goal is to train a model to detect multiple instruments and pitches playing simultaneously, the data used for training is generated by mixing a random combination of instruments and pitches.

Our idea is to perform orchestration using deep neural networks trained for classification. By generating input data that is combinations of different instruments and pitches, the network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts. Then, when a network is given a target sound as input, it will attempt to apply the same classification rules, outputing a vector of the probabilities of each sample being in the target sound. By taking the samples that have the highest probability, an orchestrated solution can be created.

We did not train our model to classify the dynamics of a sample despite TinySOL having pianissimo, mezzo forte, and fortissimo recordings for each sample. If we had, then each class would have only one data point. Instead, when the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model. If the model output a probability higher than $0.66$ for a sample, the fortissimo version of the sample was used. If it was between $0.33$ and $0.66$, then the mezzo forte version, and if less than $0.33$ the pianissimo dynamic. \textbf{\{I am not sure this hypothesis can be used accurately\}}

>>>>

In order to increase variability in our generated data, we used two methods of data augmentation.

The first one was done on the soundfile resulting from the mixture of the samples, before the mel spectrogram was taken. We performed a slight pitch shift by shifting the frequency of the signal. We did this by ... \textbf{Alejandro, can you add how you did the pitch shift (and maybe explain why changing the SR leads to a pitch shift)}In practice, the sample rate used to compute the Mel spectrogram was modified to be within $5\%$ of the actual $44100$Hz.

The second augmentation was performed on the feature matrix itself, the mel spectrogram. We chose random columns and rows of the matrix to zero out. For a given matrix, each column and each row had individually a $1\%$ chance to be set to 0, which yielded an average of $1.28$  columns and $3.45$ rows being zero-ed out. This method of data augmentation aimed to be more resilient to the possible errors of recording, due either to a setup that induces interferences and cancels specific frequencies, or to a default in the microphone that fails to record a small temporal portion.

\section{Our model}

Our idea is to perform orchestration using deep neural networks trained for classification. In our setting, each class is an instrument-note pair. Our model takes in the mel spectrogram of a soundfile that is a combination of different instruments and pitches, and attempts to identify which instrument-pitch pairs are present. The network learns how to take a sound that is a complex combination of notes and timbres and deconstruct it into its original parts. Then, when a network is given a target sound as input, it will attempt to apply the same classification rules, outputting a vector of the probabilities of each sample being in the target sound. By taking the samples that have the highest probability, an orchestrated solution can be created.

We did not train our model to classify the dynamics of a sample despite TinySOL having pianissimo, mezzo forte, and fortissimo recordings for each sample. If we had, then each class would be an instrument-pitch-dynamic tuple and would have only one data point. Instead, when the model is used for orchestration, the dynamic is determined by the probability of that sample as output by the model. If the model output a probability higher than $0.66$ for a sample, the fortissimo version of the sample was used. If it was between $0.33$ and $0.66$, then the mezzo forte version, and if less than $0.33$ the pianissimo dynamic. The idea behind this is that samples that are the most represented in the target should be the loudest in the orchestrated solution. \textbf{\{I am not sure this hypothesis can be used accurately\}}
